Intelligent Reflecting Surface-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design

### 2021-04-09 main.tex
:pencil: Update existing discussions

			\begin{remark}\label{re:reflection_coefficient}
				The element impedance $Z_l$ maps to the reflection coefficient $\phi_l$ uniquely. Since the reactance $X_l$ depends on the frequency, the reflection coefficient $\phi_l$ is also a function of frequency and cannot be designed independently at different subbands. In this paper, we assume the bandwidth is incomparable to the operating frequency such that the reflection coefficients of each IRS element are the same at all subbands. That is to say, $\phi_l$ is shared by $M$ auxiliary channels over $N$ subbands.
			\end{remark}

@@

			Similar to Algorithm~\ref{al:bcd}, we propose a Low-Complexity-BCD (LC-BCD) algorithm that iteratively update the passive beamforming $\boldsymbol{\phi}$ by Algorithm~\ref{al:lc}, the active beamforming $\boldsymbol{b}_{\mathrm{I/P},n}$, $\forall n$ by equation \eqref{eq:b}, and the waveform amplitude $\boldsymbol{s}_{\mathrm{I/P}}$ by \eqref{eq:wf}, \eqref{eq:smf} for TS or \eqref{eq:s_i}, \eqref{eq:s_p} for PS, until convergence. Similar to Proposition~\ref{pr:bcd}, strict convergence and local optimality of the LC-BCD algorithm are guaranteed if $\boldsymbol{\Phi}$ is rank-\num{1}.
		\end{subsection}
	\end{section}

@@

		We first provide numerical proof on the strict convergence (thus local optimality) of both BCD algorithms. Too see this, we define the eigen ratio as $\nu=\max_l\lambda_l(\boldsymbol{\Phi})/\sum_l\lambda_l(\boldsymbol{\Phi})$ where $\lambda_l(\boldsymbol{\Phi})$ is the $l$-th eigenvalue of $\boldsymbol{\Phi}$. $\nu=1$ means $\boldsymbol{\Phi}$ is rank-\num{1}, and we claim $\boldsymbol{\Phi}$ is approximately rank-\num{1} when $1-\nu \approx 0$. Table~\ref{ta:eigen_ratio} shows the maximum value of $1-\nu$ over all R-E samples during both BCD algorithms. With reasonable precision, we conclude that $\boldsymbol{\Phi}$ is always rank-\num{1} under different configurations for all tested channel realizations. It suggests that the rank constraint \ref{co:irs_rank} can be dropped without performance loss, $\bar{\boldsymbol{\phi}}$ can be directly obtained through eigen decomposition, and the assumption of Proposition~\ref{pr:bcd} always hold such that both BCD algorithms converge to local optimal.

@@

		Fig.~\ref{fi:re_subband} illustrates the average R-E region versus the number of subband $N$. \textit{First}, it is observed that increasing $N$ reduces the per-subband rate $R/N$ but boosts the harvested energy. The reason is that the power budget is divided into smaller portions for each subband, but the number of balanced DC terms $m_\mathrm{P}$ is increased to boost the output DC current, confirming the scaling laws in \cite{Clerckx2018b}. Sorted waveform amplitudes in Fig.~\ref{fi:waveform_subband} also confirmed that, from the perspective of WPT, a dedicated multisine waveform is unnecessary for a small $N$ but is required for a large $N$. As shown in \eqref{eq:y_I4} and \eqref{eq:y_P4}, the only difference between modulated and unmodulated waveform on $z$ exists in the fourth-order terms, where $\mathbb{E}\left\{\mathbb{A}\left\{y_{\mathrm{I}}^4(t)\right\}\right\}$ has $N^2$ monomials with a modulation gain of \num{2} and $\mathbb{A}\left\{y_{\mathrm{P}}^4(t)\right\}$ has $(2N^3+N)/3$ monomials without modulation gains. Therefore, the superposed waveform enlarges the R-E region for a sufficiently large $N$ (typically no smaller than 4). \textit{Second}, the R-E region is convex for $N = 2, 4$ and concave-convex for $N = 8, 16$. This has the consequence that PS outperforms TS for a small $N$ and is outperformed for a large $N$. When $N$ is in between, the optimal strategy is a combination of both, i.e. a time sharing between the WPT point and the saddle SWIPT point obtained by PS (as the red curve in Fig.~\ref{fi:re_subband}). Compared with the linear harvester model that requires no dedicated power waveform and always prefer PS, the rectifier nonlinearity enlarges the R-E region by favoring a different waveform and transceiving mode, both heavily depends on $N$.

@@

		The influence of the average noise power on the R-E region is investigated in Fig.~\ref{fi:re_noise}. \textit{First}, we note that the R-E region is roughly concave/convex at low/high SNR such that TS/PS are preferred correspondingly. This is because the WF strategy tends to concentrate power on few strongest subbands at low SNR. As the rate constraint $\bar{R}$ decreases, more subbands are activated to further boost the harvested DC power thanks to the frequency compensation effect and harvester nonlinearity. \textit{Second}, there exists a turning point in the R-E region especially for a low noise level ($\sigma_n^2 \le \SI{-40}{\dBm}$). The reason is that when $\bar{R}$ departs slightly from the maximum value, the algorithm tends to adjust the splitting ratio $\rho$ rather than allocating more power to the multisine waveform, since a small-amplitude multisine could be inefficient for energy maximization. As $\bar{R}$ further decreases, a modulated waveform with a very large $\rho$ could be outperformed by a superposed waveform with a smaller $\rho$, owing to the high energy efficiency of multisine. The result highlights the benefit of superposed waveform and the necessity of joint waveform and splitting ratio optimization.

@@

		The number of transmit antennas $M$ and IRS elements $L$ influence the average R-E tradeoff as revealed in Fig.~\ref{fi:re_tx} and \ref{fi:re_reflector}. \textit{First,} it is observed that adding either active or passive elements benefits both information and power transmission. This is because increasing $M$ or $L$ indeed enhances the equivalent composite channel strength. \textit{Second,} the number of active and passive elements have negligible impacts on the optimal transceiving mode (TS/PS). Unlike the noise, the channel strength influences both information and power transfer, thus varying $M$ or $L$ simultaneously scales both rate and current and preserves the general trend of the R-E boundary. \textit{Third,} the passive beamforming has a larger array gain and power scaling order than the active beamforming. This behavior is more obvious in Fig.~\ref{fi:scaling_tx} and \ref{fi:scaling_reflector} that illustrate the performance of WIT and WPT only. For the active MRT beamforming, doubling $M$ brings a \SI{3}{\dB} gain at the average output SNR, corresponding to a transmit array gain of $M$ and a doubled harvester input power. Due to the rectifier nonlinearity, the output DC current ends up with a nearly four-time (\SI{12}{\dB}) increase, validating the active scaling law in the order of $M^2$ \cite{Clerckx2016a}. On the other hand, when the IRS is very close to the transmitter or the receiver, doubling $L$ can bring an SNR boost up to \SI{6}{\dB}, corresponding to a reflect array gain of $L^2$. An interpretation is that the IRS coherently combines multiple signal copies with a receive array gain $L$, then performs an equal gain reflection with a transmit array gain $L$, achieving a total gain of $L^2$. As a consequence, doubling IRS elements brings a four-fold boost on the received signal power that further amplifies the harvested DC current by \num{16} times (\SI{24}{\dB}), suggesting a passive scaling law in the order of $L^4$. Besides, by considering harvester nonlinearity, the proposed design brings a nearly \SI{20}{\dB} increase in the output DC current over the conventional linear perspective. These results emphasize the effectiveness of passive beamforming and the importance of modeling rectifier nonlinearity in the SWIPT system design. \textit{Fourth,} small-scale SWIPT is dominated by the active beamforming while large-scale SWIPT relies more on the passive beamforming. As shown in Fig.~\ref{fi:path_loss}, in our setup, the direct path loss $\Lambda_\mathrm{D}$ is around \SI{60}{\dB} while the auxiliary path loss product $\Lambda_{\mathrm{I}}\Lambda_R$ is over \SI{100}{\dB}. Hence, to fully exploit the R-E advantage of passive beamforming, a large number of reflecting elements are required to enhance the auxiliary channel strength. Another explanation is that the RF-chains of active antennas provide design flexibility in the frequency domain, while the frequency-flat characteristic of the IRS restricts its benefit for broadband transmission, especially when $L$ is small.

@@

		Fig.~\ref{fi:re_irs_1mhz} and \ref{fi:re_irs_10mhz} explore the average R-E region under different IRS configuration for narrowband transmission ($B=\SI{1}{\MHz}$) and broadband transmission ($B=\SI{10}{\MHz}$). The rate-dependent adaptive IRS strategy adjusts the passive beamforming for each point on the R-E boundary, while the WIT/WPT-based non-adaptive IRS schemes only optimize the phase shifts for information and power purposes (correspond to the points on the $x$ and $y$ axes, respectively). To gain some insight into the IRS behavior, we compare the results above to no IRS and an ideal Frequency-Selective (FS) IRS. In contrast to Remark~\ref{re:irs_frequency_flat}, each element in the ideal FS IRS is assumed to have an independent reflection coefficient for each subband, accounting for $NL$ variables in total. When there is only one transmit antenna, each ideal IRS element simply aligns the AP-IRS-UE channel to the AP-UE channel at all subbands, namely
		\begin{equation}
			\theta_{l,n}^{\star}=e^{j\arg{\left(h_{\mathrm{D},n}/h_{\mathrm{I},n,l}h_{\mathrm{R},n,l}\right)}},\quad \forall n,l.
		\end{equation}
		\textit{First}, it is observed that the assistance of the IRS effectively enlarges the achievable R-E region for both narrowband and broadband transmissions. \textit{Second}, the performance of all four IRS strategies coincide with each other for narrowband SWIPT. It suggests that the optimal passive beamforming obtained for any R-E point is also optimal for the whole R-E region. Therefore, for a given channel, we can fix the active and passive beamforming, and only adapt the waveform and splitting ratio to achieve different R-E performance. This is because when the bandwidth is small, all channels are approximately flat such that the optimal ideal reflection coefficients at all subbands are approximately the same. Also, the subchannel tradeoff in Remark~\ref{re:irs_frequency_flat} does not exist as the auxiliary channel can be simultaneously maximized at different frequencies. \textit{Third}, a joint optimization of waveform, active and passive beamforming, and splitting ratio is necessary to maximize the R-E tradeoff for broadband SWIPT. Recall Remark~\ref{re:irs_subchannel_alignment} that WIT and WPT prefer different subchannel strength distribution. By adaptive passive beamforming, the channel strength can either be concentrated in few strongest subbands to enhance the rate at a low SNR, or be spread evenly over all subbands to boost the output DC power. The advantage of such a smart channel control is prominent for broadband SWIPT where the strength can vary significantly for different subbands.

>>>>

			\begin{remark}\label{re:reflection_coefficient}
				The element impedance $Z_l$ maps to the reflection coefficient $\phi_l$ uniquely. Since the reactance $X_l$ depends on the frequency, the reflection coefficient $\phi_l$ is also a function of frequency and cannot be designed independently at different subbands. In this paper, we assume the bandwidth is incomparable to the operating frequency such that the reflection coefficient of each IRS element is the same at all subbands. That is to say, $\phi_l$ is shared by $M$ auxiliary channels over $N$ subbands.
			\end{remark}

			\begin{remark}\label{re:subband_tradeoff}
				The cascaded channel response varies at different frequencies. Since we assume the reflection coefficient to be frequency-flat, there exists a tradeoff for auxiliary channel design in the frequency domain. Moreover, each IRS element simultaneously assists all transmit antennas such that the reflection coefficient is shared by $M$ cascaded channels over $N$ subbands.
			\end{remark}

@@

			Similar to Algorithm~\ref{al:bcd}, we propose a Low-Complexity-BCD (LC-BCD) algorithm that iteratively update the passive beamforming $\boldsymbol{\phi}$ by Algorithm~\ref{al:lc}, the active beamforming $\boldsymbol{b}_{\mathrm{I/P},n}$, $\forall n$ by equation \eqref{eq:b}, and the waveform amplitude $\boldsymbol{s}_{\mathrm{I/P}}$ by \eqref{eq:wf}, \eqref{eq:smf} for TS or \eqref{eq:s_i}, \eqref{eq:s_p} for PS, until convergence. Similar to Proposition~\ref{pr:bcd}, strict convergence and local optimality of the LC-BCD algorithm are guaranteed if $\boldsymbol{\Phi}$ is rank-\num{1}. Details are omitted here for simplicity.
		\end{subsection}
	\end{section}

@@

		We first provide numerical proof on the strict convergence (thus local optimality) of both BCD algorithms. Too see this, we define the eigen ratio as $\nu=\max_l\lambda_l(\boldsymbol{\Phi})/\sum_l\lambda_l(\boldsymbol{\Phi})$ where $\lambda_l(\boldsymbol{\Phi})$ is the $l$-th eigenvalue of $\boldsymbol{\Phi}$. Since $\nu=1$ means $\boldsymbol{\Phi}$ is rank-\num{1}, we claim $\boldsymbol{\Phi}$ is approximately rank-\num{1} when $1-\nu \approx 0$. Table~\ref{ta:eigen_ratio} shows the maximum value of $1-\nu$ over all R-E samples during both BCD algorithms. With reasonable precision, we conclude that $\boldsymbol{\Phi}$ is always rank-\num{1} under different configurations for all tested channel realizations. It suggests that the rank constraint \ref{co:irs_rank} can be relaxed without performance loss, $\bar{\boldsymbol{\phi}}$ can be directly obtained through eigen decomposition, and the assumption of Proposition~\ref{pr:bcd} always hold such that both BCD algorithms converge to local optimal.

@@

		Fig.~\ref{fi:re_subband} illustrates the average R-E region versus the number of subband $N$. \textit{First,} it is observed that increasing $N$ reduces the per-subband rate $R/N$ but boosts the harvested energy. It is because each subband receives less power but more balanced DC terms are introduced to boost the harvested energy. On the other hand, Fig.~\ref{fi:waveform_subband} sorts the modulated/multisine amplitude $\boldsymbol{s}_{\mathrm{I/P}}$ for WPT in descending order. It demonstrated that a dedicated multisine waveform is unnecessary for a small $N$ but is required for a large $N$. This observation origins from the rectifier nonlinearity. Although both waveforms have equivalent second-order DC terms \eqref{eq:y_I2} and \eqref{eq:y_P2}, for the fourth-order DC terms \eqref{eq:y_I4} and \eqref{eq:y_P4}, the modulated waveform has $N^2$ monomials with a modulation gain of \num{2} while the multisine has $(2N^3+N)/3$ monomials. Hence, the benefit of multisine outstands for a sufficiently large $N$. \textit{Second,} the R-E region is convex for $N = 2, 4$ and concave-convex for $N = 8, 16$. This has the consequence that PS outperforms TS for a small $N$ and is outperformed for a large $N$. When $N$ is in between, the optimal strategy is a combination of both, i.e. a time sharing between the WPT point and the saddle SWIPT point obtained by PS (as the red curve in Fig.~\ref{fi:re_subband}). Compared with the linear harvester model that requires no dedicated power waveform and always prefer PS, the rectifier nonlinearity enlarges the R-E region by favoring a different waveform and transceiving mode, both heavily depends on the number of subbands.

@@

		The average noise power influences the R-E region as in Fig.~\ref{fi:re_noise}. \textit{First,} we note that the R-E region is roughly concave/convex at low/high SNR such that TS/PS are preferred correspondingly. At low SNR, the power is allocated to the modulated waveform on few strongest subbands to achieve a high rate. As the rate constraint $\bar{R}$ decreases, Algorithm~\ref{al:gp} activates more subbands that further boosts the harvested DC power due to frequency balancing and harvester nonlinearity. \textit{Second,} there exists a turning point in the R-E region especially for a low noise level ($\sigma_n^2 \le \SI{-40}{\dBm}$). The reason is that when $\bar{R}$ departs slightly from the maximum value, the algorithm tends to adjust the splitting ratio $\rho$ rather than allocating more power to the multisine waveform, since a small-amplitude multisine could be inefficient for energy purpose. As $\bar{R}$ further decreases, due to the advantage of multisine, a modulated waveform with a very large $\rho$ could be outperformed by a superposed waveform with a smaller $\rho$. The result proves the benefit of superposed waveform and the necessity of joint waveform and splitting ratio optimization.

@@

		The impacts of the number of transmit antennas $M$ and IRS elements $L$ on the R-E behavior are revealed in Fig.~\ref{fi:re_tx} and \ref{fi:re_reflector}. \textit{First,} adding either active or passive elements enhances the equivalent channel strength to boost information and power transfer. Interestingly, it has negligible influence on the optimal transceiving strategy because both rate and current are amplified that preserves the general trend of the R-E boundary. \textit{Second,} the passive beamforming has a larger array gain and power scaling order than the active beamforming. This behavior is more obvious in Fig.~\ref{fi:scaling_tx} and \ref{fi:scaling_reflector} for the case of WIT and WPT, where the Adaptive Single Sinewave (ASS) allocates all power to the sine wave on the strongest subband and is optimal for the linear harvester model \cite{Clerckx2016a}. For the MRT active precoder, doubling $M$ brings a \SI{3}{\dB} gain at the output SNR that corresponds to a transmit array gain of $M$ and a doubled harvester input power. Thanks to the rectenna nonlinearity, the output DC current ends up with a nearly four-time (\SI{12}{\dB}) increase that validates the active scaling law in the order of $M^2$ \cite{Clerckx2016a}. On the other hand, when the IRS is very close to the AP, doubling $L$ increases the output SNR up to \SI{6}{\dB} that implies a reflect array gain of $L^2$. An interpretation is that the IRS coherently combines the incoming signal with a receive array gain $L$, then performs an equal gain reflection with a transmit array gain $L$. Hence, doubling IRS elements brings a four-fold increase on the received signal power that further amplifies the harvested DC current by \num{16} times (\SI{24}{\dB}), corresponding to a passive scaling law in the order of $L^4$. Compared with active antennas, the IRS achieves higher array gain and power scaling order by passive elements with frequency-dependent reflection, but a very large $L$ is required to compensate the double fading of the auxiliary link. Moreover, the ASS strategy exploit no harvester nonlinearity and ends up with a nearly \SI{20}{\dB} gap with SMF and GP. These observations demonstrate the R-E benefit of passive beamforming and emphasize the importance of accounting the harvester nonlinearity in the passive beamforming design.

@@

		Fig.~\ref{fi:re_irs_1mhz} and \ref{fi:re_irs_10mhz} explore the R-E region with different IRS strategies for narrowband and broadband SWIPT. The ideal Frequency-Selective (FS) IRS assumes the reflection coefficient of each element is independent and controllable at different frequencies. The adaptive IRS adjusts the passive beamforming for different rate constraints by Algorithm~\ref{al:sca}. The WIT/WPT-optimized IRS is retrieved by Algorithm~\ref{al:lc} then fixed for the whole R-E region (i.e. the boundary is obtained by varying the waveform amplitude and splitting ratio). \textit{First,} random IRS and no IRS perform worse than other schemes since no passive beamforming is exploited. Their achievable R-E regions coincide because the antenna mode reflection of the random IRS is cancelled out after averaging over channel realizations. \textit{Second,} the performance of non-random IRS are similar when the bandwidth is small, while the adaptive IRS outperforms the WIT/WPT-optimized IRS but is worse than the ideal FS IRS when the bandwidth is large. In the former case, the subband responses are close to each other such that the tradeoff in Remark~\ref{re:subband_tradeoff} becomes unobvious and the auxiliary channels can be roughly maximized at all subbands. It suggests that for narrowband SWIPT, the optimal passive beamforming of any R-E point is roughly optimal for the whole R-E region, which can be approximated in the closed form (by aligning the auxiliary channel with the direct channel) when a single transmit antenna is used at the AP. On the other hand, the optimal reflection coefficient varies at different R-E points for broadband SWIPT, since the information and power transfer prefer different channel strength distribution in the frequency domain. By adaptive passive beamforming, the channel strength can be either amplified at few strongest subbands to enhance the rate at low SNR, or spread evenly to boost the output DC power.

		% The rate-dependent adaptive IRS strategy adjusts the passive beamforming for each point on the R-E boundary, while the WIT/WPT-based non-adaptive IRS schemes only optimize the phase shifts for information and power purposes (correspond to the points on the $x$ and $y$ axes, respectively). To gain some insight into the IRS behavior, we compare the results above to no IRS and an ideal Frequency-Selective (FS) IRS.

		% \textit{First,} it is observed that the assistance of the IRS effectively enlarges the achievable R-E region for both narrowband and broadband transmissions.
		% performance of all four IRS strategies coincide with each other for narrowband SWIPT. It suggests that the optimal passive beamforming obtained for any R-E point is also optimal for the whole R-E region. Therefore, for a given channel, we can fix the active and passive beamforming, and only adapt the waveform and splitting ratio to achieve different R-E performance. This is because when the bandwidth is small, all channels are approximately flat such that the optimal ideal reflection coefficients at all subbands are approximately the same. Also, the subchannel tradeoff in Remark~\ref{re:irs_frequency_flat} does not exist as the auxiliary channel can be simultaneously maximized at different frequencies. \textit{Third,} a joint optimization of waveform, active and passive beamforming, and splitting ratio is necessary to maximize the R-E tradeoff for broadband SWIPT. Recall Remark~\ref{re:irs_subchannel_alignment} that WIT and WPT prefer different subchannel strength distribution. By adaptive passive beamforming, the channel strength can either be concentrated in few strongest subbands to enhance the rate at a low SNR, or be spread evenly over all subbands to boost the output DC power. The advantage of such a smart channel control is prominent for broadband SWIPT where the strength can vary significantly for different subbands.
